\documentclass[oneside]{amsart}
\input{preamble}
\input{notation}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\addbibresource{pwg.bib}

\newcommand{\pShv}{\prescript{p}{}{\Shv}}
\newcommand{\plMod}{\prescript{p}{}{\lMod}}
\DeclareMathOperator{\supp}{supp}
\newcommand{\ptau}{\prescript{p}{}{\tau}}

\title{Stabilization of cohomology on the Ran space}
\author{Micha{\l} Mruga{\l}a}

\begin{document}
\maketitle

\setcounter{section}{-1}
\section{Notation}
\begin{itemize}
	\item We often abbreviate $\Ran X$ to just $\Ran$.
	\item We write $\mathring{\CF}^{I}$ for $\ins_I^{!}(\CF)|_{\mathring{X}^{I}}$. Similarly for $I$ replaced by $n$.
	\item If we apply multiple restrictions in a row as in the previous item, we !-pushforward back to $\Ran$ between each !-pullback.
\end{itemize}

\section{The Goal}
The goal of this talk is to prove two theorems which appeared in the last talk. They concern conditions for (complexes of) sheaves on the Ran space to behave well with respect to Verdier duality. 

\begin{definition}
	Let $a,b$ be integers $\ge 0$. We say that a sheaf $\CF\in \Shv^{!}(\Ran X)$ \defn{satisfies} $P(a,b)$ if there exists an integer $m\ge 0$ such that for any finite set $I$ with $|I|>m$
	\[
		\mathring{\CF}^{I}\in \pShv(\mathring{X}^{I})^{\le -a-b|I|}.
	\] 
\end{definition}

Recall that we are using the perverse $t$-structure. Such a $t$-structure for a variety $V$ over $k$ is given by
\[
\pShv(V)^{\le 0} = \left\{ \CF\in \Shv(V) \middle| \forall i\in \Z,\dim\supp H^{i}(\CF) \le -i \right\}. 
\] 
We will often use the simple fact:
\begin{proposition}\label{prop:ptoc}
	Let $\CF\in \pShv(V)^{\le n}$, then $\CF\in \Shv(V)^{\le n}$.
\end{proposition}
\begin{proof}
	The case $n=0$ is clear. The general case follows by shifting.
\end{proof}

We want to prove the following theorems:
\begin{theorem}\label{thm:vd}
	Let $X\in \Sch$ be a proper, smooth curve and $\CF\in \Shv^{!}(\Ran X)$. Suppose that for every integer $a\ge 0$, $\CF$ satisfies $P(a,1)$. Then the map
	\[
	C^{*}_c(\Ran, \VD\CF) \to C^{*}_c(\Ran, \CF)^{\vee}
	\] 
	is an isomorphism.
\end{theorem}

\begin{theorem}\label{thm:prod}
	Let $X$ be a smooth curve and $\Lambda$ have finite cohomological dimension. Let $\CF_1,\CF_2\in \Shv^{-}(\Ran X)$ satisfy $P(a,1)$ for every integer $a\ge 1$. Assume furthermore that $\CF_1,\CF_2$ have compact cohomology sheaves. Then the map
	\[
		(\VD\CF_1) \boxtimes (\VD\CF_2) \to \VD(\CF_1\boxtimes\CF_2)
	\] 
	is an isomorphism.
\end{theorem}

These results are used in the paper in two places:
\begin{enumerate}[(i)]
	\item \Cref{thm:prod} is used to prove that given a certain set of conditions on the curve $X$, the ring of coefficients $\Lambda$, and a factorization algebra $\CG$ on $\Ran X$ the Verdier dual is also a factorization algebra.
	\item Recall from Antoine's talk that we can reduce the proof of the cohomological product formula to constructing an isomorphism
		\[
		C^{*}_c(\Ran X, \CB) \to C^{*}_c(\Ran X, \CA)^{\vee}.
		\] 
		This is almost exactly the statement of \Cref{thm:vd} if $\CB\isoto \VD\CA$. However, $\CA,\CB$ don't satisfy the cohomological estimate in \Cref{thm:vd}, so we need to replace them by their reduced versions $\CA\red,\CB\red$ which do satisfy them. But I want to highlight that \Cref{thm:vd} is really at the core of the main proof in \cite{abott}.
\end{enumerate}

\section{Proof of the First Theorem}
\textbf{Step 1: Reduction to the $\le 0$ case.}

We need to show that for every integer $k\ge 0$ the map
\[
	C^{*}_c(\Ran, \VD\CF)^{\le k} \to \left(C^{*}_c(\Ran, \CF)^{\vee}\right)^{\le k}
\] 
is an isomorphism. To make book-keeping easier we reduce to the $k=0$ case, which we can do since if $\CF$ satisfies $P(a,1)$ so does $\CF[i]$ for $i\ge 0$.

\textbf{Step 2: Reduction to the truncated Ran space.}

For any $n\ge 0$ there is a commutative diagram
\[
\begin{tikzcd}
	C^{*}_c(\Ran, \VD\CF)^{\le 0} \ar[r] \ar[d] & (C^{*}_c(\Ran,\CF)^{\vee})^{\le 0} \ar[d] \\
	C^{*}_c(\Ran^{\le n},\VD\CF^{\le n})^{\le 0} \ar[r] & (C^{*}_c(\Ran^{\le n},\CF^{\le n})^{\vee})^{\le 0}
\end{tikzcd}
\] 
If the left and right arrows are isomorphisms, it suffices to prove that the bottom arrow is an isomorphism. We defer the proofs for the left and right arrows to the next two sections.

\textbf{Step 3: Proof for the truncated Ran space.}
Since $\Ran^{\le n}$ is finitary pseudo-proper this follows from \cite[Corollary~7.5.6]{abott} which was covered in Andrea's talk.

\section{The Right Arrow}
We prove the following (stronger) statement.

\begin{proposition}\label{prop:ti}
	Let $X$ be a smooth curve and $\CF\in \Shv^{!}(\Ran X)$. Suppose $\CF$ satisfies $P(3,0)$, then the map
	\[
	C^{*}_c\left( \Ran^{\le n},\CF^{\le n} \right) ^{\ge 0} \to C^{*}_c\left( \Ran, \CF \right) ^{\ge 0}
	\] 
	is an isomorphism for $n\gg 0$.
\end{proposition}
\textbf{Step 1: Reduction to the truncated Ran space.}

Since
\begin{equation}\label{eq:colim}
C^{*}_c(\Ran,\CF) = \colim_{k\in \N} C^{*}_c(\Ran^{\le k},\CF^{\le k}),
\end{equation} 
there exists $M\in \Z$ such that for all $m\ge M$ the map
\[
C_c^{*}(\Ran^{\le m},\CF^{\le m}) \to C^{*}_c(\Ran,\CF)
\] 
is an isomorphism. Hence it suffices to show
\begin{lemma}\label{prop:tit}
	Let $X$ be a smooth curve and $\CF\in \Shv^{!}(\Ran X)$. Suppose $\CF$ satisfies $P(3,0)$, then the map
	\[
	C^{*}_c(\Ran^{\le n},\CF^{\le n})^{\ge 0} \to C^{*}_c(\Ran^{\le m},\CF^{\le m})^{\ge 0}
	\] 
	is an isomorphism for $n\gg 0$ and $m\ge n$.
\end{lemma}

\textbf{Step 2: Reduction to the calculation of the cofiber.}

It suffices to handle the $m=n+1$ case. Let
\[
D = \CoFib\left(  C^{*}_c(\Ran^{\le n},\CF^{\le n}) \to C^{*}_c(\Ran^{\le n+1},\CF^{\le n+1})
\right). 
\] 
Using the long exact sequence in cohomology, it suffices to show that
\[
D\in \lMod^{\le -2}.
\] 

\begin{lemma}\label{lmm:cofibcoh}
	$D$ is canonically isomorphic to $C^{*}(\mathring{X}^{n}, \mathring{\CF}^{n})_{\Sigma_{n+1}}$.
\end{lemma}
Here $\Sigma_{n+1}$ is the symmetric group on $n+1$ elements.

Assuming \Cref{lmm:cofib}, it suffices to show that
\[
	C^{*}(\mathring{X}^{n},\mathring{\CF}^{n}) \in \lMod^{\le 2}.
\] 
since taking coinvariants is right-exact. By \Cref{prop:ptoc} it suffices to show that $C^{*}(\mathring{X}^{n},\mathring{\CF}^{n})\in \plMod^{\le 2}$.

We can factor the structure morphism of $\mathring{X}^{n}$ as
\[
\begin{tikzcd}[column sep=huge]
	\mathring{X}^{n} \ar[r,"p","\text{affine}"'] & X \ar[r,"q","\text{smooth curve}"'] & \Spec k.
\end{tikzcd}
\] 
Since $p$ is affine, $p_*$ is right $t-$exact for the perverse $t$-structure \cite[Theorem~4.1.1]{BBD}. Since $q$ is quasiprojective, $q_*$ is cohomologically bounded by $\le 1$ for the perverse $t$-structure \cite[4.2.4]{BBD}.

\textbf{Step 3: Calculation of the cofiber.}
To prove \Cref{lmm:cofibcoh} we first calculate
\[
	\CoFib(\ins^{\le n}_!\CF^{\le n}\to \ins^{\le n+1}_!\CF^{\le n+1})\in \Shv^{!}(\Ran).
\] 
Write $\ins^{n,n+1}$ for the inclusion $\Ran^{\le n}\to \Ran^{\le n+1}$. Then the cofiber is isomorphic to
\[
\ins^{\le n+1}_! \CoFib(\ins^{n,n+1}\CF^{\le n}\to \CF^{\le n+1}),
\] 
so we need to compute the cofiber on $\Ran^{\le n+1}$.

Since $\ins^{n,n+1}$ is finitary pseudo-proper and injective, $\ins^{n,n+1}_!$ defines an equivalence of categories between $\Shv^{!}(\Ran^{\le n})$ and sheaves on its image by \cite[Lemma~7.4.11(d)]{abott}. Hence the cofiber is supported on the complement
\[
	(\Ran^{\le n+1}-\Ran^{\le n})\subset \Ran^{\le n+1}.
\] 

Now we give a characterization of this complement as almost a scheme. The morphism $\ins_{n+1}$ factors through
\[
	\ins_{n+1} /\Sigma_{n+1}:X^{n+1} /\Sigma_{n+1} \to \Ran.
\] 
Writing $j_{n+1}$ for the inclusion $\mathring{X}^{n+1}\to X^{n+1}$ we get an inclusion
\[
	j_{n+1} /\Sigma_{n+1}:\mathring{X}^{n+1} /\Sigma_{n+1} \to X^{n+1} /\Sigma_{n+1}.
\] 
$\ins_{n+1} /\Sigma_{n+1}$ factors through $\Ran^{\le n+1}$ and restricts to an isomorphism
\[
	\mathring{X}^{n+1}/\Sigma_{n+1} \simto \Ran^{\le n+1} - \Ran^{\le n}.
\] 

\begin{lemma}\label{lmm:cofib}
	Let $\CF\in \Shv^{!}(\Ran)$, then
	\[
	\CoFib\left( \ins_!^{\le n}\CF^{\le n} \to \ins_!^{\le n+1}\CF^{\le n+1} \right) 
	\] 
	is canonically isomorphic to
	\[
		(\ins_{n+1} /\Sigma_{n+1})_! \circ (j_{n+1} /\Sigma_{n+1})_* (\mathring{\CF}^{n+1} /\Sigma_{n+1}).
	\] 
\end{lemma}
\begin{proof}
	This follows from the localization exact sequence. (I don't fully understand why we are guaranteed the existence of this exact sequence in this case.)
\end{proof}
\begin{proof}[Proof of \Cref{lmm:cofibcoh}]
	This follows from applying $C^{*}_c$. (Note that the paper doesn't ask for $X$ to be proper, but I don't know how we can get the structure morphism from $X^{n+1} /\Sigma_{n+1}$ to define the same $*$ and $!$-pushforwards otherwise.)
\end{proof}

\section{The Left Arrow}
We prove the following (stronger statement):
\begin{proposition}\label{prop:dualti}
	Let $X$ be a smooth, proper curve and $\CF\in \Shv^{!}(\Ran X)$. Suppose that $\CF$ satisfies $P(2,1)$. Then the map
	\[
	C^{*}_c(\Ran, \VD\CF)^{\le 0} \to C^{*}_c \left( \Ran^{\le n}, \VD \CF^{\le n} \right) ^{\le 0}
	\] 
	is an isomorphism for $n\gg 0$.
\end{proposition}
In the proof of \Cref{prop:ti} we used the fact that $C^{*}_c(\Ran,-)$ commutes with colimits, but it does not commute with limits, which is a problem since
\begin{equation}\label{eq:lim}
\VD\CF \isoto \lim_n \ins^{\le n}_! \left( \VD \CF^{\le n} \right) .
\end{equation} 
However, for $k\ge 0$ the functors $C^{*}_c(\Ran^{\le k},-)$ and $\ins^{\le k,!}$ \emph{do} commute with limits, so we want to reduce to the $C^{*}_c(\Ran^{\le k},-^{\le k})$ case first.

\textbf{Step 1: Reduction to the truncated Ran space.}

Using \eqref{eq:colim} it suffices to show that for all $k\ge 0$ 
\[
C^{*}_c(\Ran^{\le k},(\VD\CF)^{\le k})^{\le 0} \to C^{*}_c(\Ran^{\le k},(\ins^{\le n}_!(\VD\CF^{\le n}))^{\le k})^{\le 0}
\] 
is an isomorphism.

\textbf{Step 2: Commuting the limit.}
Now we plug in \eqref{eq:lim} and use the fact that $C^{*}_c(\Ran^{\le k},-^{\le k})$ commutes with limits. Hence it suffices to show:
\begin{lemma}
	Let $X$ be a smooth, proper curve and $\CF\in \Shv^{!}(\Ran X)$ be a sheaf satisfying $P(2,1)$, then for any $k\ge 0$ the map
	\[
		C^{*}_c \left( \Ran^{\le k}, \left( \ins^{\le n}_! \left( \VD\CF^{\le n} \right) \right) ^{\le k} \right) ^{\le 0} \to C^{*}_c \left( \Ran^{\le k}, \left( \ins^{\le n-1}_! \left( \VD\CF^{\le n-1} \right) \right) ^{\le k} \right) ^{\le 0} 
	\] 
	for $n\gg 0$.
\end{lemma}

\textbf{Step 3: Reduction to $\mathring{X}^{k}$.}
To prove the lemma it suffices to show the the fiber
\[
\CG = \Fib \left( \left( \ins^{\le n}_! \left( \VD\CF^{\le n} \right)  \right) ^{\le k} \to  \left( \ins^{\le n}_! \left( \VD\CF^{\le n-1} \right)  \right) ^{\le k}\right)
\] 
satisfies $C^{*}_c(\Ran^{\le k},\CG)\in \lMod^{\ge 2}$.

Using the filtration
\[
\Ran^{\le 1}\subset \Ran^{\le 2}\subset \dots\subset \Ran^{\le k}
\] 
we see that it suffices to show that for all $k'\le k$
\[
\CoFib \left( C^{*}_c(\Ran^{\le k'-1},\CG^{\le k'-1}) \to C^{*}_c(\Ran^{\le k'},\CG^{\le k'}) \right) \in \lMod^{\ge 2}.
\] 
Note that we can just take $k'=k$ since $k$ is arbitrary $\ge 0$.

By \Cref{lmm:cofibcoh} this cofiber is isomorphic to
\[
	C^{*}\left( \mathring{X}^{k},\mathring{\CG}^{k} \right)_{\Sigma_k}. 
\] 
Since $C^{*}(\mathring{X}^{k},\mathring{-}^{k})$ is exact, this is isomorphic to
\[
	\Fib\left( C^{*}\left( \mathring{X}^{k},\mathring{(\VD\CF^{\le n})}^{k} \right) \to C^{*}\left( \mathring{X}^{k}, \mathring{(\VD\CF^{\le n-1})}^{k} \right)  \right)_{\Sigma_k}.
\] 
Since for any sheaf $\CH\in \Shv^{!}(\mathring{X}^{k} /\Sigma_k)$ the norm map
\[
	C^{*}(\mathring{X}^{k},\mathring{\CH}^{k})_{\Sigma_k} \to C^{*}(\mathring{X}^{k},\mathring{\CH}^{k})^{\Sigma_k}
\] 
is an isomorphism, taking (co)invariants is exact, so it suffices to show that the fiber is concentrated in degrees $\ge 2$.

\textbf{Step 4: Calculation of the fiber.}

Since $C^{*}(\mathring{X}^{k},\mathring{-}^{k})$ is left exact, we are reduced to computing the fiber
\[
\Fib \left( \ins^{\le n}_!(\VD\CF^{\le n}) \to \ins^{\le n-1}_!(\VD\CF^{\le n-1}) \right) 
\] 
on $\Ran$. Since $\ins^{n-1,n}$ is finitary pseudo-proper it commutes with Verdier duality, so
\begin{align*}
	\ins_!^{\le n-1} \VD\CF^{\le n-1} &\isoto \ins_!^{\le n}\ins_!^{n-1,n} \VD\CF^{\le n-1} \\
					    &\isoto \ins_!^{\le n} \VD\ins_!^{n-1,n}\CF^{\le n-1}.
\end{align*}
Hence we can pull $\Fib$ through $\ins_!^{\le n}\VD$ and need to calculate
\[
\CoFib\left( \ins_!^{n-1,n}\CF^{\le n-1} \to \CF^{\le n} \right) ,
\] 
but by \Cref{lmm:cofib} we know that it is isomorphic to
\[
(\ins_{n} /\Sigma_{n})_! \circ (j_{n} /\Sigma_{n})_* (\mathring{\CF}^{n} /\Sigma_{n})
\] 
on $\Ran^{\le n}$. Plugging this back in and unfolding compositions we find that the original fiber is
\[
	C^{*}\left( \mathring{X}^{k}, j_k^{!}\ins_k^{!}\ins_{n,!}\left( \VD j_{n,*} j_n^{!}\CF^{n} \right) \right)^{\Sigma_n}. 
\] 
And again by exactness of invariants on $\mathring{X}^{k}$, we need to show that the complex before taking invariants is concentrated in degrees $\ge 2$.

Hence it suffices to prove:
\begin{lemma}\label{lmm:xkcoh}
	Let $\CH\in \pShv(X^{n})^{\le -n-2}$, then
	\[
		C^{*}\left( \mathring{X}^{k},j_k^{!}\ins_k^{!}\ins_{n,!}(\VD\CH) \right) \in \lMod^{\ge 2}.
	\] 
\end{lemma}

\textbf{Step 5: Reduction to schemes.}
Recall from my first talk that if
\[
X^{k,n}(S) = \left\{ (x_1,\dots,x_k,y_1,\dots,y_n)\in X^{k}(S)\times X^{n}(S) \middle| \{x_1,\dots,x_k\} = \{y_1\dots,y_n\}  \right\} 
\] 
and we write
\[
\begin{tikzcd}
	X^{k} & X^{k,n} \ar[l,"q_k"'] \ar[r,"q_n"] & X^{n}
\end{tikzcd}
\] 
for the projections, then
\[
\ins_k^{!}\circ \ins_{n,!} \isoto q_{k,!}\circ q_n^{!}.
\] 
So we are reduced to proving that
\[
	C^{*}\left( \mathring{X}^{k},j_k^{!}q_{k,!}q_n^{!}(\VD\CH) \right) \in \lMod^{\ge 2}.
\] 

\textbf{Step 6: The scheme case.}

Here we will use $a$ to generally refer to structure morphisms. We will also make use of the following Cartesian square
\[
\begin{tikzcd}
	\mathring{X}^{k,n} \ar[r,"j_{k,n}"] \ar[d,"\mathring{q}_k"] \ar[dr,"\wt{q}_k"] & X^{k,n} \ar[d,"q_k"] \\
	\mathring{X}^{k} \ar[r,"j_k"] & X_k.
\end{tikzcd}
\] 
By properness of $q_k$ we can use proper base change to get
\begin{align*}
	a^{k}_*j_k^{!}q_{k,!}q_n^{!}\VD\CH &\isoto a^{k}_*\mathring{q}_{k,!}j_{k,n}^{!}q_n^{!}\VD\CH \\
					   &\isoto \mathring{a}^{k,n}_* \wt{q}_n^{!}\VD\CH \\
					   &\isoto a^{n}_* \wt{q}_{n,*}\wt{q}_n^{!}\VD\CH.
\end{align*} 
Now we can use standard six functor formalism non-sense to rewrite
\[
	\wt{q}_n^{!}\VD\CH \isoto (\wt{q}_n^{*}\VD\CH) \otimes^{!} \omega.
\] 
Applying the projection formula we find that our desired expression is
\[
	C^{*}\left( X^{n},(\VD\CH)\otimes^{!}(\wt{q}_{n,*}\omega) \right) 
\] 
Since $\wt{q}_{n,*}\omega$ is compact this is isomorphic to
\[
	\Hom(\CH,\wt{q}_{n,*}\omega).
\] 
Since $\dim X^{n}=n$ and $\wt{q}_n$ is quasifinite, $\wt{q}_{n,*}\omega\in \pShv(X^{n})^{\ge -n}$. And now the conclusion follows from our assumption on $\CH$.

\section{A Variant of the Theorem}
For the proof of the cohomological product formula we will need the following variant:
\begin{theorem}
	Let $X$ be a proper, smooth curve and $X'\subset X$ be an open subscheme. Let $\CF'\in \Shv^{!}(\Ran X')$ satisfy $P(a,1)$ for every integer $a\ge 0$. For $\CF \defeq \bb{j}_{!}\CF'$ the map
	\[
	C^{*}_c(\Ran, \VD\CF) \to C^{*}_c(\Ran,\CF)^{\vee}
	\] 
	is an isomorphism.
\end{theorem}
Here $\bb{j}$ is the morphism $\Ran X'\to \Ran X$.

Following the proof of \Cref{thm:vd}, and using the fact that \Cref{prop:ti} is true for non-proper curves, we are left to deal with the left arrow
\[
C^{*}_c(\Ran, \VD\CF)^{\le 0} \to C_c^{*}(\Ran^{\le n}, \VD\CF^{\le n})^{\le 0}.
\] 
Now we use the isomorphisms
\[
	\VD\circ \bb{j}_! \isoto \bb{j}_* \circ \VD,\qquad \VD \circ \bb{j}^{\le n}_! \isoto \bb{j}^{\le n}_*\circ \VD
\] 
to reduce to a lemma analogous to \Cref{prop:dualti}. Then the proof is nearly identical.

\section{Computing Duality at Finite Levels}
To prove \Cref{thm:prod} we first need to show that we can compute $\VD$ on $\Ran^{\le n}$ for sufficiently large $n$.
\begin{proposition}\label{prop:vdtrunc}
	Let $X$ be a smooth curve and $\CF\in \Shv^{!}(\Ran X)$. Suppose that $\CF$ satisfies $P(2,1)$. Then for any $\CI$ and $n\gg 0$ the map
	\[
		\ptau^{\le 0}(\VD\CF)^{\CI} \to \ptau^{\le 0}(\VD\CF^{\le n})^{\CI}
	\] 
	is an isomorphism.
\end{proposition}

\textbf{Step 1: Reduction to fiber computation.}

Without loss of generality we can replace $X^{\CI}$ by $X^{k}$ for $k\ge 0$. Since $\ins_k^{!}$ commutes with limits, we can replace the left hand side by
\[
\ptau^{\le 0} \left( \lim_{n'} (\VD\CF^{\le n'})^{k} \right) .
\] 
As before we can reduce to showing that
\[
\left( \Fib \left( \VD\CF^{\le n} \to \VD\CF^{\le n-1} \right)  \right) ^{k} \in \pShv(X^{k})^{\ge 2}.
\] 

\textbf{Step 2: Fiber computation.}

As before, we compute the fiber to be
\[
	\left( \ins_k^{!}\ins_{n,!}\VD j_{n,*}\mathring{\CF}^{n} \right) ^{\Sigma_n}
\] 
and we can drop the $\Sigma_n$ invariants. Let
\[
	\CG \defeq j_{n,*}\mathring{\CF}^{n} \in \pShv(X^{n})^{\le -n-2}.
\] 

\textbf{Step 3: Reduction to scheme case.}

As before rewrite
\[
\ins_k^{!}\ins_{n,!}\VD\CG \isoto q_{k,!}q_n^{!}\VD\CG.
\] 

\textbf{Step 4: The scheme case.}

Since $q_k,q_n$ are both finite, $q_{k,!}$ is $t$-exact and $q_n^{!} $ is left $t$-exact for the perverse $t$-structure. Hence it suffices to show that $\VD\CG\in \pShv(X^{n})^{\ge 2}$ which is standard.

\section{\texorpdfstring{Proof of \Cref{thm:prod}}{Proof of Product Theorem}}

Note that in this section we will be working with the perverse $t$-structure.

\textbf{Step 1: Initial reductions.}

We need to show that for any $m_1,m_2,k\ge 0$ we have an isomorphism
\[
\left( \left( \VD\CF_1 \boxtimes \VD\CF_2 \right) ^{m_1,m_2} \right) ^{\le k} \to \left( \left( \VD\left( \CF_1\boxtimes\CF_2 \right)  \right) ^{m_1,m_2} \right) ^{\le k}
\] 
is an isomorphism. As before we can reduce to the $k=0$ case.

\textbf{Step 2: Reduction to the truncated case.}

For $n\ge 0$ we have a commutative diagram
\[
\begin{tikzcd}
	\left( \left( \VD\CF_1\boxtimes\VD\CF_2 \right) ^{m_1,m_2} \right) ^{\le 0} \ar[r] \ar[d] & \left( \left( \VD\left( \CF_1\boxtimes\CF_2 \right)  \right) ^{m_1,m_2} \right) ^{\le 0} \ar[d] \\
	\left( \left( \VD\CF_1^{\le n}\boxtimes\VD\CF_2^{\le n} \right) ^{m_1,m_2} \right) ^{\le 0} \ar[r]  & \left( \left( \VD\left( \CF_1^{\le n}\boxtimes\CF_2^{\le n} \right)  \right) ^{m_1,m_2} \right) ^{\le 0} 
\end{tikzcd}
\] 

We claim that the left and right arrows are isomorphisms for $n\gg 0$, then the proof is reduced to showing that the top arrow is an isomorphism. We postpone the proof of this to the next section.

\textbf{Step 3: Reduction to the scheme case.}

We prove a stronger statement:
\begin{lemma}
	For $n\ge 0$ the map
	\[
	\left( \ins^{\le n}\times \ins^{\le n} \right) _! \left( \VD\CF_1^{\le n}\boxtimes\VD\CF_2^{\le n} \right) \to \left( \ins^{\le n}\times \ins^{\le n} \right) _! \VD \left( \CF_1^{\le n}\boxtimes\CF_2^{\le n} \right) 
	\] 
	is an isomorphism.
\end{lemma}
The left hand side is isomorphic to
\[
	\lim_{\CI_1,\CI_2\in \Fins[n]} \left( \ins_{\CI_1}\times \ins_{\CI_2} \right) _! \left( \VD\CF_1^{\CI_1}\boxtimes\VD\CF_2^{\CI_2} \right) .
\] 
Similarly, the right hand side is isomorphic to
\[
	\lim_{\CI_1,\CI_2\in \Fins[n]} \left( \ins_{\CI}\times \ins_{\CI_2} \right) _! \left( \VD \left( \CF_1^{\CI_1}\boxtimes\CF_2^{\CI_2} \right)  \right) .
\] 
So it suffices to show that for every $\CI_1,\CI_2\in \Fins$ the map
\[
\VD\CF_1^{\CI_1}\boxtimes \CF_2^{\CI_2} \to \VD\left( \CF_1^{\CI_1}\boxtimes\CF_2^{\CI_2} \right) .
\] 

\textbf{Step 4: The scheme case.}
As $\CF_i^{\CI_i}$ are bounded above with compact cohomology sheaves we are done by \cite[Lemma~7.3.2]{abott}.

\section{Side Maps II}
We focus on the left map. The argument for the right map is almost identical, but requires a variant of \Cref{prop:vdtrunc} to be proved first.

\textbf{Step 1: Reduction to cohomological bound.}

Let
\[
\CG_i \defeq (\VD\CF_i)^{m_i},\qquad \CG_i(n) \defeq \left( \VD\CF_i^{\le n} \right) ^{m_i}
\] 
and
\[
\CE_i(n) \defeq \CoFib(\CG_i \to \CG_i(n)).
\] 
We have an exact triangle
\[
\begin{tikzcd}
	\CE_1(n) \boxtimes \CG_2 \ar[r] & \CoFib\left( \CG_1\boxtimes\CG_2 \to \CG_1(n)\boxtimes\CG_2(n) \right) \ar[r] & \CG_1(n) \boxtimes \CE_2(n).
\end{tikzcd}
\] 
By \Cref{prop:vdtrunc} for $n\gg 0$ we have $\CE_i(n)\in \pShv(X^{m_i})^{\ge 1}$. Note that we will certainly need to show that $\CG_i$ and $\CG_i(n)$ are bounded below for the perverse $t$-structure. Assuming this, and using the fact that $\Lambda$ has finite Tor dimension, we can certainly show that the cofiber is bounded below for the perverse $t$-structure. I'm not quite sure how to show it can get large enough, though.

\begin{remark}
	In our case we are fine, since we can improve the bound in \Cref{prop:vdtrunc} to be linear in $n$. At the last step of that proof we can show that $\VD\CG\in \pShv(X^{n})^{\ge n+2}$. However, this argument fails in the $D$-module case.
\end{remark}

\textbf{Step 2: Lower bound.}

By \Cref{prop:vdtrunc}, to show that $\CG_i$ is bounded below it suffices to show $\CG_i(n)$ is bounded below for $n\gg 0$, so we've reduced to the $\CG_i(n)$ case.

But
\[
	\ins^{\le n}_!\left( \VD\CF_i^{\le n} \right) \isoto \lim_{\CI\in \Fins[n]} \ins_{\CI,!}\left( \VD\CF_i^{\CI} \right) .
\] 
Since this limit is finite, it suffices to show that for each $\CI\in \Fins[n]$ 
\[
	(\VD\CF_i^{\CI})^{m_i} \in \Shv^{+}(X^{m_i}).
\] 

By assumption $\CF_i^{\CI}$ is bounded above, so $\VD\CF_i^{\CI}$ is bounded below. Now since
\[
\ins_{m_i}^{!}\circ \ins_{\CI,!} \isoto q_{\CI,!}\circ q_{m_i}^{!}
\] 
is bounded below, we are done.

\printbibliography

\end{document}
